---
title: "Where to Eat during the AGA 2015 PDT Conference?"
subtitle: "Data Analysis and Visualization of Yelp Ratings in Nashville, TN"
author: "FI Consulting"
date: "July 12-15, 2015"
output: 
  ioslides_presentation:
    logo: ../img/logo_slides.png
    transition: faster
    widescreen: no
---

<style>
slide{background-color:white;}
td{font-size:14px;}
article{color:black;}

.title-slide hgroup h1{font-size:45px;color:#1142F5;}
.title-slide hgroup h2{font-size:28px;color:#C5A619;}
table{margin:auto;}

.smaller_font p,li{font-size:18px;}

</style>


```{r setup, include=FALSE}
library(knitr)
library(leaflet)
library(ggplot2)
library(ggthemes)
library(DT)
library(scales)
library(png)
library(jpeg)
library(grid)
library(dplyr)

opts_chunk$set(echo = FALSE, results='asis', size="small", fig.align = 'center')
theme_set(theme_fivethirtyeight())

# load data
load('../data/clean/nash_restaurants.RData')
load('../data/clean/nash_hot_chicken.RData')
load('../data/clean/nash_music.RData')

# Normalize ratings
nash <- mutate(nash,
               yelp_qrt = cut(rating, 
                              breaks = quantile(rating, seq(0,1,length.out = 5)), 
                              include.lowest = T,
                              labels = 1:4))

# list of categories
categ <-  data.frame(
  cat = unlist(lapply(nash$categories, strsplit, split = ':')),
  cat_cd = unlist(lapply(nash$categories_cd, strsplit, split = ':')),
  stringsAsFactors = F) %>% 
  unique() %>% 
  arrange(rank(cat_cd))

# center coord for leaflet maps
nash.view = list(lat =36.15970,
                 lng = -86.76060,
                 zoom = 12)
# Conference location and Icon
agaIcon <- icons(
  iconUrl = "../img/aga.png",
  iconWidth = 422/5.626667, iconHeight = 233/5.626667)

# functions for best in category
top_hood <- function(categ.best){
  datatable(
    nash %>% 
      filter(grepl(categ.best, categories_cd)
             & !is.na(neighborhood)) %>%
      group_by(neighborhood) %>% 
      summarize(count = n(),
                dist = round(weighted.mean(dist, review_count),2),
                rating = round(weighted.mean(rating, review_count),2),
                review_count = sum(review_count)) %>% 
      arrange(-rating*review_count) %>% 
      filter(row_number()<=5),
    colnames = c('Neighborhood','Restaurants', 'Distance','Rating', 'Reviews'),
    class = 'row-border stripe compact order-column hover',
    options=list(pageLength = 5, dom = 't'),
    rownames=FALSE)
}

top_restr <- function(categ.best){
  datatable(
    nash %>% 
      filter(grepl(categ.best,categories_cd)) %>% 
      top_n(20,rating*review_count) %>% 
      arrange(-rating*review_count) %>%
      mutate(Rank = row_number()) %>% 
      select(Rank, name, neighborhood, categories ,rating, review_count,dist),
    colnames = c('Rank','Name','Neighborhood', 'Category', 'Rating', 'Reviews','Distance'),
    class = 'row-border stripe compact order-column hover',
    options=list(pageLength = 5, dom = 'tip'),
    rownames=FALSE)
}

top_map <- function(categ.best, size_adj=1){
  
  df <- nash %>% 
    filter(grepl(categ.best, categories_cd)) %>%
    top_n(20,rating*review_count) %>% 
    arrange(-rating*review_count) %>%
    mutate(Rank = row_number(),
           popup = paste0(
             name, 
             '<BR>Rank: ', Rank, 
             '<BR>Neighborhood: ', neighborhood, 
             '<BR>Categories: ', categories, 
             '<BR>Rating: ', rating, 
             '<BR>Reviews: ', review_count, 
             '<BR>Distance: ', dist
           )) %>% 
    select(popup, lat, lng, review_count, Rank)
  
  # Create  color palette
  pal <- colorNumeric(
    palette = c('forestgreen', 'yellow', 'tomato'),
    domain = df$Rank
  )
  
  leaflet(data = df, 
          width = 800, height = 450) %>% 
    addProviderTiles("CartoDB.Positron") %>%
    setView(lng = nash.view$lng, lat = nash.view$lat, zoom = nash.view$zoom) %>% 
    # addMarkers(popup = ~popup, lng = ~lng, lat = ~lat) %>% 
    addCircleMarkers(
      popup = ~popup,
      lng = ~lng, lat = ~lat,
      radius = ~sqrt(review_count)*size_adj,
      color = ~pal(Rank),
      stroke = FALSE, fillOpacity = 0.80) %>%     
    addMarkers(lng = -86.69255, lat = 36.21149, 
               icon = agaIcon,
               popup = 'AGA Conference<BR>Gaylord Opryland Hotel & Convention Center')
}


```



## Welcome 2015 AGA PDT Attendees

```{r nash_img, fig.width=0.70*727/72, fig.height=0.70*257/72}
img <- readJPEG('../img/nashville_guitar.jpg')
grid.raster(img)
```



```{r logo, fig.width=2, fig.height=1.5}
img <- readPNG('../img/aga.png')
grid.raster(img)
```


## Yelp Ratings

```{r yelp_dist, fig.width=6, fig.height=3.5}
ggplot(nash, aes(x=rating)) + 
  ggtitle('Number of restaurants by Yelp Rating') +
  geom_histogram(binwidth = 0.50, fill='firebrick', color='white') +
  scale_x_continuous(limits=c(1, 5)) +
  scale_y_continuous(labels = comma)

kable(t(summary(nash$rating)),format = 'html', align='c')

```



----

__Restaurant Categories__  
```{r cloud, fig.width=6, fig.height=6}
img <- readPNG('../img/wordcloud.png')
grid.raster(img)
```



## {.smaller_font}
__What is Nashville <span style='color:red;font-weight: bold'>Hot Chicken</span>?__  

* Marinated in seasoning, floured, fried, and sauced with <span style='color:red;'>__cayenne pepper__</span> paste 
* Served atop slices of white bread with pickle chips

```{r chic_img, fig.width=0.85*500/72, fig.height=0.85*323/72}
img <- readJPEG('../img/nashville_hot_chicken.jpg')
grid.raster(img)
```

----

**Top Restaurants Serving Nashville Hot Chicken**  

```{r}

datatable(
    hot.chic %>% 
      top_n(20,-rank) %>% 
      arrange(rank) %>%
      select(rank, name ,rating, review_count,dist, snippet_text),
    colnames = c('Rank','Name', 'Rating', 'Reviews','Distance','Reviews Snippet'),
    class = 'row-border stripe compact order-column hover',
    options=list(pageLength = 5, dom = 'tip'),
    rownames=FALSE)


```

##{.smaller_font}

**Top Places to Find Nashville's Hot Chicken**

```{r hot_chic_map}

df = hot.chic %>%
            top_n(20, -rank) %>% 
            mutate(popup = paste0(
                     name, 
                     '<BR>Rank: ', rank, 
                     '<BR>Categories: ', categories, 
                     '<BR>Rating: ', rating, 
                     '<BR>Reviews: ', comma(review_count))) %>% 
            select(popup, lat, lng, rating, matches, rank)

# Create  color palette
pal <- colorNumeric(
  palette = c('forestgreen', 'yellow', 'tomato'),
  domain = df$rank
)

leaflet(data = df, width = 800, height = 500) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng = nash.view$lng, lat = nash.view$lat+0.01, zoom = nash.view$zoom) %>% 
  addCircleMarkers(
    popup = ~popup,
    lng = ~lng, lat = ~lat,
    radius = ~sqrt(matches)*2.5,
    color = ~pal(rank),
    stroke = FALSE, fillOpacity = 0.90) %>% 
  addMarkers(lng = -86.69255, lat = 36.21149, 
             icon = agaIcon,
             popup = 'AGA Conference<BR>Gaylord Opryland Hotel & Convention Center')




```


## Explore Nashville's Music Scene {.smaller_font}

```{r music_img, fig.width=0.50*1140/72, fig.height=0.50*260/72}
img <- readJPEG('../img/Nashville_Broadway.jpg')
grid.raster(img)
```

**Top Neighborhoods** 
```{r top_music_hoods}

  datatable(
    nash.music %>% 
      filter(!is.na(neighborhood)) %>%
      group_by(neighborhood) %>% 
      summarize(count = n(),
                dist = round(weighted.mean(dist, review_count),2),
                rating = round(weighted.mean(rating, review_count),2),
                review_count = sum(review_count)) %>% 
      arrange(-rating*review_count) %>% 
      filter(row_number()<=5),
    colnames = c('Neighborhood','Venues', 'Distance','Rating', 'Reviews'),
    class = 'row-border stripe compact order-column hover',
    options=list(pageLength = 5, dom = 't'),
    rownames=FALSE)


```

## Top Music Venues
```{r top_music_venues}

datatable(
  nash.music %>% 
    filter(!grepl('cinema',categories_cd)) %>% 
    top_n(20,rating*review_count) %>% 
    arrange(-rating*review_count) %>%
    mutate(categories = gsub('Music Venues:','', categories),
           categories = gsub(':',', ', categories),
           Rank = row_number()) %>% 
    select(Rank, name, categories ,rating, review_count,dist,snippet_text),
  colnames = c('Rank','Name', 'Category', 'Rating', 'Reviews','Distance','Review Snippet'),
  class = 'row-border stripe compact order-column hover',
  options=list(pageLength = 5, dom = 'tip'),
  rownames=FALSE)



```


## {.smaller_font}

__Where to Find Nashville's Best Music Venues__  
```{r music_map}

df = nash.music %>%
  filter(!grepl('cinema',categories_cd)) %>% 
  top_n(20, -rank) %>% 
  mutate(rank = row_number(),
         categories = gsub('Music Venues:','', categories),
           categories = gsub(':',', ', categories),
         popup = paste0(
           name, 
           '<BR>Rank: ', rank, 
           '<BR>Categories: ', categories, 
           '<BR>Neighborhood: ', neighborhood, 
           '<BR>Rating: ', rating, 
           '<BR>Reviews: ', comma(review_count))) %>% 
  select(popup, lat, lng, rating, review_count, rank)

# Create  color palette
pal <- colorNumeric(
  palette = c('forestgreen', 'yellow', 'tomato'),
  domain = df$rank
)

leaflet(data = df, width = 800, height = 500) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng = nash.view$lng, lat = nash.view$lat+0.01, zoom = nash.view$zoom) %>% 
  addCircleMarkers(
    popup = ~popup,
    lng = ~lng, lat = ~lat,
    radius = ~sqrt(review_count)*0.75,
    color = ~pal(rank),
    stroke = FALSE, fillOpacity = 0.90) %>% 
  addMarkers(lng = -86.69255, lat = 36.21149, 
             icon = agaIcon,
             popup = 'AGA Conference<BR>Gaylord Opryland Hotel & Convention Center')




```


